---
title: "Application CELLector to pancancer collection (Ciriello)"
author: "Lucia Trastulla"
date: "14/12/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(chatgpt, lib.loc = "~/R/x86_64-pc-linux-gnu-library/4.1/")
```

### Introduction

The data is obtained from a collaboration with Giovanni Ciriello. The collection includes multiple data sets and is divided per missense and truncating variants. We use the latest CELLector version that includes the partitioned one.

```{r, message=FALSE, warning=FALSE}
library(devtools)
if (!require("CELLector", character.only = TRUE)) {
  install_github("Francescojm/CELLector")
}
if (!require("ggplot2", character.only = TRUE)) {
  install.packages("ggplot2")
}
if (!require("tidyverse", character.only = TRUE)) {
  install.packages("tidyverse")
}
library(CELLector)
library(ggplot2)
library(tidyverse)
```

### Load data

```{r,  fig.height = 9, fig.width = 6}

input_fold <- "/group/iorio/Datasets/tumour_binary_Ciriello/"
meta_table <- read_tsv(sprintf("%sall_merged_metadata.tsv", input_fold), 
                       show_col_types = FALSE)
object_to_load <- list.files(input_fold, pattern = ".rds")

obj <- list()
for (id_name in seq_len(length(object_to_load))) {
  
  obj[[id_name]] <- readRDS(paste0(input_fold, object_to_load[id_name]))
  obj[[id_name]]$M$M$total <- obj[[id_name]]$M$M$missense + obj[[id_name]]$M$M$truncating
  obj[[id_name]]$M$M$total <- apply(obj[[id_name]]$M$M$total > 0, c(1,2), as.numeric)
  meta_tmp <- meta_table[match(colnames(obj[[id_name]]$M$M$total), meta_table$sample), ]
  obj[[id_name]]$sample_info <- meta_tmp
}

tot_sample_info <- lapply(obj, function(x) x$sample_info) %>%
  dplyr::bind_rows()

t_factor <- tot_sample_info %>% 
  dplyr::group_by(Tissue) %>%
  dplyr::summarise(count = n()) %>%
  dplyr::arrange(desc(count)) %>% 
  dplyr::pull(Tissue)

tot_sample_info <- tot_sample_info %>% 
  dplyr::mutate(Tissue = factor(Tissue, levels = as.character(t_factor)))

ggplot(tot_sample_info, aes(x = Tissue, fill = Cohort)) + 
  geom_bar() + 
  theme_bw() +
  theme(axis.title.y = element_blank()) + 
  coord_flip()

```

### Store data separately for each cancer type

```{r}

extract_cancer <- function(obj_cohort, 
                           var_name = "Tissue"){
  
  BEM <- obj_cohort$M$M$total 
  type_name <- obj_cohort$sample_info %>% dplyr::pull(get(var_name)) %>% unique()
  BEM_split <- purrr::map(type_name, function(x) BEM[,obj_cohort$sample_info$Tissue == x])
  names(BEM_split) <- type_name
  
  return(BEM_split)    
}

build_BEM_cancer_type <- function(obj_split, tissue){
  
  obj_single_tissue <- obj_split %>% keep(names(.) %in% c(tissue))
  common_mut <- purrr::map(obj_single_tissue, rownames) %>% 
    purrr::flatten_chr() %>%
    unique()

  # take union and assign to 0 for non-available ones (CORRECT?)
  obj_harmonized_single_tissue <- purrr::map(obj_single_tissue, function(x)
    if (any(!common_mut %in% rownames(x))) {
      rbind(x, matrix(0, 
                      nrow = length(setdiff(common_mut, rownames(x))), 
                      ncol = ncol(x),
                      dimnames = list(setdiff(common_mut, rownames(x)), colnames(x))))
    }else{x})

  obj_harmonized_single_tissue <- purrr::map(obj_harmonized_single_tissue, 
                                           function(x) x[common_mut,])
  BEM_single_tissue <- purrr::reduce(obj_harmonized_single_tissue, cbind)
  meta_table_tissue <- meta_table[match(colnames(BEM_single_tissue), meta_table$sample),]
  
  return(list(BEM = BEM_single_tissue, 
              meta = meta_table_tissue))
  
}

obj_split_tissue <- purrr::map(obj, function(x) extract_cancer(x)) %>% 
  purrr::flatten()
tissues_name <- unique(names(obj_split_tissue))

BEM_all <- purrr::map(tissues_name, function(x) 
  build_BEM_cancer_type(obj_split_tissue, x))

names(BEM_all) <- tissues_name

```

### Apply CELLector

```{r}

create_CSSp_groups <- function(CSS_p){
  
  id_samples <- str_split(CSS_p$Points, pattern = "[,]")
  id_sign <- purrr::map2(.x = id_samples, .y = CSS_p$Signature, 
                         .f = ~ rep(.y, length(.x)))
  df <- data.frame(sample = flatten_chr(id_samples), 
                   CELLectorP_Sign = flatten_chr(id_sign))
  return(df)
}

CSSp_group_complete <- list()
CSS <- list()

for (id_t in seq_len(length(tissues_name))) {

  cancer_type <- tissues_name[id_t]
  print(paste("######", cancer_type, "######"))
  
  minGlobSupp_par <- 0.01
  #n_samples <- ncol(BEM_all[[cancer_type]]$BEM)
  #if (floor(n_samples*0.01) < 20) {
  #   minGlobSupp_par <- 20/n_samples
  #}
  
  CSS[[id_t]] <- CELLector.Build_Search_Space_Partitioned(
    ctumours = t(BEM_all[[cancer_type]]$BEM), 
    cancerType = cancer_type, 
    mutOnly = TRUE, 
    UD_genomics = TRUE, 
    verbose = FALSE, 
    minGlobSupp = minGlobSupp_par)

  print(CELLector.visualiseSearchingSpace_sunBurst(searchSpace = CSS[[id_t]]$hierarchical))

  CSSp_group <- create_CSSp_groups(CSS_p = CSS[[id_t]]$partitioned)
  meta_tissue <- BEM_all[[cancer_type]]$meta
  CSSp_group_complete[[id_t]] <- full_join(CSSp_group, meta_tissue) %>% 
    dplyr::mutate(CSS_minGlobSupp =  minGlobSupp_par)
  
}

```

### Are there enrichments in subtypes?

### Download survival info, differences?

### Download CCL data and project, differences on essentiality?
